### 2024-11-14 paper/paper.qmd
Added sub-sections

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(ggplot2)
library(dplyr)
library(knitr)
library(rstanarm)
library(arrow)
library(data.table)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

For this study, we utilize data from Project Hammer, a Canadian initiative designed to monitor grocery prices across major retailers in Canada. The Project Hammer dataset was collected from eight prominent Canadian grocery vendors—Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods—between February 28, 2024, and the latest available data load in the database [ citeProjectHammerData]. This dataset enables us to investigate price fluctuations, identify pricing trends, and explore potential competitive or collusive patterns within the Canadian grocery sector.
Project Hammer aims to support regulatory efforts by providing data on price transparency, making it suitable for academic research and potential legal analysis. Following data processing and cleaning procedures outlined in Section 3.2, the dataset was structured to allow an analysis of historical price changes and vendor-specific price patterns.
The dataset contains 1,996,969 rows and 5 columns, with variables capturing product (product name), vendor_name (name of the grocery retailer), price_current (current product price at the time of recording), price_old (previous recorded price for the product), and price_difference (difference between the current and old prices). Price data is recorded in Canadian dollars and captures a broad range of grocery items from various categories, including fresh produce, dairy, pantry staples, and household items.
To ensure the quality and consistency of the data, we focused on removing missing values and imputing outlier prices when necessary. For this analysis, we excluded entries with extreme price fluctuations beyond three standard deviations, as these could represent temporary discounts or data recording anomalies. This cleaning process provides a balanced and accurate reflection of grocery pricing, allowing for robust statistical analysis on vendor-specific trends and average price comparisons across product types.
By narrowing the scope to recent data, Project Hammer's dataset captures real-time fluctuations in grocery pricing, providing timely insights into potential market dynamics and consumer price impacts. The analysis allows for future investigation of temporal price trends, vendor-based price differences, and, more broadly, consumer affordability in a market facing rising cost-of-living pressures.

## Measurement

## Outcome variables

## Predictor variables

# Model

Our modeling approach seeks to explore and quantify the relationship between previous grocery prices, vendor identities, and price differences in the current prices observed within Canadian grocery stores. This analysis employs a Bayesian linear regression model implemented via the `stan
_glm` function in the `rstanarm` package to examine how factors such as historical prices, vendor differences, and observed price changes impact current prices for various grocery items.

In this model, `price_current` serves as the response variable, while `price_old`, `vendor_name`, and `price_difference` act as predictor variables. The linear regression model assumes a Gaussian distribution for the response variable `price_current`, allowing for a straightforward interpretation of the estimated parameters.

## Model set-up

The model includes the following predictor variables:

- Previous Price (`price_old`): The price of the product in a previous time period.

- Vendor (`vendor_name`): A categorical variable representing the grocery store chain selling the product, capturing vendor-specific pricing differences.

- Price Difference (`price_difference`): The difference between the current and previous price, which may indicate market or vendor-specific pricing adjustments.

The model can be represented mathematically as follows:

\begin{align*}
    y_i \mid \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
    \mu_i &= \beta_0 + \beta_1 \cdot \text{Previous Price}_i + \beta_2 \cdot \text{Vendor}_i + \beta_3 \cdot \text{Price Difference}_i \\
    \epsilon_i &\sim \text{Normal}(0, \sigma^2)
\end{align*}

\textbf{Where:}
\begin{itemize}
    \item $\beta_0$ is the intercept term, representing the baseline estimate of $price_{current}$
    \item $\beta_1$, $\beta_2$, and $\beta_3$ are the coefficients representing the effects of $price_{old}$, $vendor_{name}$ and $price_{difference}$ on $price_{current}$.
    \item $\sigma^2$represents the variance of the error term, capturing unexplained variability in current prices.
\end{itemize}

The model is executed in \texttt{R} [@citeR] using the \texttt{rstanarm} package [@rstanarm],  with priors set to regularize the estimates and prevent overfitting. Specifically, we use a normal prior for the coefficients, centered at zero with moderate variance, to ensure stable estimates without overly restrictive assumptions.

### Model justification

The selection of a linear regression model for this analysis is based on the continuous nature of `price_current`, which allows us to quantify how previous pricing, vendor, and price changes influence current pricing. Existing economic theories suggest that previous prices and historical pricing patterns are often predictive of current pricing, especially in competitive retail markets where price sensitivity and vendor-specific strategies play significant roles. The inclusion of `price_old` as a predictor aligns with time-series economic models where past data points inform future values.

Additionally, vendor differences, represented by `vendor_name`, capture potential competitive dynamics in the Canadian grocery market. Vendors may adopt distinct pricing strategies or respond differently to market conditions, and the inclusion of this categorical variable enables an analysis of these patterns across various major grocery chains. The `price_difference` variable, which represents recent price changes, provides an understanding of market adjustments and reflects fluctuations potentially influenced by external factors, such as supply chain constraints or inflation.

The Bayesian approach was selected for its flexibility in incorporating prior information and its robustness in handling uncertainty within small or moderate sample sizes. Furthermore, Bayesian methods allow us to capture the posterior distributions of the parameters, which can be used to evaluate the strength and credibility of each predictor's effect on current prices. This approach also aligns with the objectives of Project Hammer by facilitating a transparent and probabilistically rigorous analysis of grocery pricing trends.

In summary, this model helps explain the primary factors of current grocery prices in Canadian stores, accounting for both historical prices and vendor effects. By utilizing Bayesian framework, this analysis provides a clear quantification of the influence of each predictor while allowing for robust estimation and interpretability in the context of economic and competitive pricing strategies.

# Results

The results of our analysis of the Project Hammer dataset shows significant findings into pricing trends across major Canadian grocery vendors. By examining factors such as prior pricing (`price_old`), vendor identities (`vendor_name`), and recent price changes (`price_difference`), we assess how these variables affect current prices (`price_current`). The following sections summarize key findings, supported by visualizations that highlight trends and relationships within the data.

## Vendor-Specific Price Trends

We begin by examining average current prices (`price_current`) across different grocery vendors to identify potential pricing variations in the Canadian grocery market.

```{r}
#| label: fig-avg-curr-price
#| fig-cap: Average Current Price by Vendor. This bar plot shows the average price of products sold by each grocery vendor in Canadian dollars (CAD). Variations in pricing suggest differences in market positioning or pricing strategies among vendors, with some retailers consistently offering higher or lower prices across product categories.
#| echo: false
#| warning: false
#| message: false

#### Read data ####
# Load the cleaned dataset
analysis_data <- read_csv("C:/Users/eliza/Downloads/hammer-3-compressed/cleaned_price_changes.csv", show_col_types = FALSE)

# Generate plot for Average Current Price by Vendor
vendor_price_plot <- analysis_data %>%
  group_by(vendor_name) %>%
  summarize(avg_price_current = mean(price_current, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(vendor_name, avg_price_current), y = avg_price_current)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Average Current Price by Vendor",
    x = "Vendor",
    y = "Average Current Price (CAD)"
  ) +
  theme_minimal()

# Display the plot
vendor_price_plot

```
@fig-avg-curr-price displays the average current prices across vendors, revealing notable pricing differences. Metro and Galleria, for example, tend to have higher average prices, while No Frills and Walmart are on the lower end. These discrepancies suggest differing market positioning, with some vendors potentially catering to more price-sensitive customers.

**Implications for Consumers**: Budget-conscious consumers might prefer Walmart or No Frills for lower average prices, while those seeking specific product selections or brands might choose Metro or Galleria despite higher prices.

## Price Difference Analysis
We further analyze price_difference, which captures the change between the current price (`price_current`) and the previous price (`price_old`). This variable provides findings into recent price adjustments across vendors.

```{r}
#| label: fig-avg-price-diff
#| fig-cap: Average Price Difference by Vendor. This bar plot shows the average price change per vendor, highlighting retailers with the most significant pricing adjustments. Positive differences suggest recent price increases, while negative differences indicate reductions, reflecting possible discounting or competitive pricing strategies.
#| echo: false
#| warning: false
#| message: false

# Average price difference by vendor
price_diff_plot <- analysis_data %>%
  group_by(vendor_name) %>%
  summarize(avg_price_difference = mean(price_difference, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(vendor_name, avg_price_difference), y = avg_price_difference)) +
  geom_bar(stat = "identity", fill = "coral") +
  coord_flip() +
  labs(
    title = "Average Price Difference by Vendor",
    x = "Vendor",
    y = "Average Price Difference (CAD)"
  ) +
  theme_minimal()
price_diff_plot

```
@fig-avg-price-diff illustrates the average price differences by vendor. Retailers like Walmart and Galleria exhibit negative average price differences, suggesting a tendency to reduce prices, likely through competitive discounts. On the other hand, Loblaw shows relatively stable or small positive differences, suggesting less frequent or minor price adjustments.

**Competitive Strategy**: Vendors with negative average differences may be aiming to attract customers through regular discounts, while vendors with stable or positive differences might rely on consistent pricing for brand loyalty.

## Relationship Between Previous and Current Prices
To understand price consistency and elasticity, we assess the relationship between previous prices (`price_old`) and current prices (`price_current`).

```{r}
#| label: fig-past-curr
#| fig-cap: Relationship Between Previous and Current Prices. This scatter plot with a linear regression line shows the correlation between previous and current prices. A positive relationship would indicate that higher previous prices predict higher current prices, implying potential price stability or gradual price changes in the grocery market.
#| echo: false
#| warning: false
#| message: false

# Scatter plot of previous price vs. current price
price_relationship_plot <- ggplot(analysis_data, aes(x = price_old, y = price_current)) +
  geom_point(alpha = 0.4, color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Relationship Between Previous and Current Prices",
    x = "Previous Price (CAD)",
    y = "Current Price (CAD)"
  ) +
  theme_minimal()
price_relationship_plot

```
@fig-past-curr shows a strong positive relationship between previous prices and current prices. This pattern indicates that higher past prices are generally associated with higher current prices, implying stability in grocery pricing over time.

**Price Elasticity Findings**: The linear trend suggests low price elasticity, where previous prices predict current prices with relative stability. This may reflect supply chain consistency and retailers’ cautious approach to pricing changes.

## Distribution of Price Differences
To further explore pricing dynamics, we examine the distribution of `price_difference`, capturing both price increases and decreases across the dataset.

```{r}
#| label: fig-dist-price
#| fig-cap: Relationship Between Previous and Current Prices. This scatter plot with a linear regression line shows the correlation between previous and current prices. A positive relationship would indicate that higher previous prices predict higher current prices, implying potential price stability or gradual price changes in the grocery market.
#| echo: false
#| warning: false
#| message: false

# Calculate mean and standard deviation of price_difference
mean_diff <- mean(analysis_data$price_difference, na.rm = TRUE)
sd_diff <- sd(analysis_data$price_difference, na.rm = TRUE)

# Filter out extreme values beyond 3 standard deviations
filtered_data <- analysis_data %>%
  filter(price_difference >= (mean_diff - 3 * sd_diff),
         price_difference <= (mean_diff + 3 * sd_diff))

# Plot histogram with filtered data
price_difference_hist <- ggplot(filtered_data, aes(x = price_difference)) +
  geom_histogram(bins = 30, fill = "lightgreen", color = "black") +
  labs(
    title = "Distribution of Price Differences (Filtered)",
    x = "Price Difference (CAD)",
    y = "Frequency"
  ) +
  theme_minimal()
price_difference_hist

```
@fig-dist-price (filtered to remove extreme values), shows that most price changes are small, with values clustered around zero. Both positive and negative price differences are present, indicating a balance between price increases and decreases across vendors.

**Market Stability**: The concentration around zero suggests a generally stable market, with moderate pricing adjustments. While some vendors increase prices, others offer discounts, contributing to a balanced market.

# Discussion

## Correlation vs. Causation
This analysis utilizes Bayesian regression model to identify factors correlated with current grocery prices, including prior prices (`price_old`), vendor identity (`vendor_name`), and recent price differences (`price_difference`). While the model shows that historical prices and vendor-specific effects are strongly associated with current pricing, it’s essential to interpret these relationships cautiously in terms of causation.

The correlation between previous and current prices likely reflects price continuity rather than a causal mechanism. For example, stable pricing over time may arise from vendor policies or industry norms rather than from any inherent property of the product or market structure. Similarly, vendor identity correlates with price differences, but this does not imply that being a particular vendor causes higher or lower prices. Instead, vendor-specific prices likely reflect underlying operational costs, competitive positioning, or branding strategies. Caution is warranted in inferring causation, as unobserved factors—such as cost structures or consumer loyalty—may influence both vendor identity and pricing patterns.

The presence of both price increases and decreases (reflected in `price_difference`) further highlights the dynamic nature of grocery pricing. While these changes are correlated with the final prices observed, it would be speculative to conclude that specific price fluctuations directly cause current price levels. External factors like inflation, seasonal demand, and supply chain disruptions may drive these observed correlations, underscoring the need for a cautious interpretation of causality.

## Missing Data

The Project Hammer dataset contains some limitations in data coverage, which may impact the completeness and accuracy of our findings. Specifically, the data does not include smaller or regional grocery stores, which may have distinct pricing strategies and could offer lower or more competitive prices in certain areas. The absence of these smaller players may lead to an incomplete picture of the overall grocery market, as price variations across different types of stores are not captured.

Additionally, missing data on certain product categories or regions within vendors limits the granularity of the analysis. For example, without category-specific data (e.g., dairy, produce, or packaged goods), it’s challenging to examine whether price dynamics vary significantly across product types. This lack of detail may obscure category-level trends, which could be particularly relevant for staple items that heavily influence household spending. Future data collection efforts could focus on capturing a wider range of store types and product details to provide a more representative dataset.

Finally, missing information on factors like seasonal influences, promotions, or regional economic conditions means that some relevant variables are absent from the model. These factors could help explain variations in grocery pricing, particularly for products sensitive to external factors (e.g., fresh produce affected by seasonality). Addressing these data gaps could significantly improve the model’s ability to capture the full range of influences on grocery prices.

## Sources of Bias

The Project Hammer dataset and the corresponding analysis may contain several sources of bias that could affect the interpretation of results. One potential source of bias stems from the focus on major grocery chains, which may not accurately represent pricing dynamics across smaller, independent stores. Since major vendors typically have broader operational reach and standardized pricing strategies, the findings may be biased toward the practices of these larger corporations, potentially overestimating price stability or ignoring competitive pricing seen in smaller markets.

Another possible bias comes from the data’s retrospective nature, focusing on historical pricing as a predictor of current prices. This approach may inherently favor price continuity and overlook the role of short-term market disruptions, promotions, or external shocks. For instance, a focus on historical prices could understate the impact of sudden events (like supply chain breakdowns or economic crises) that may cause sharp price changes. This form of selection bias could lead to an overly conservative estimate of price volatility in the model.

Furthermore, vendor-specific effects could introduce bias if certain vendors follow distinct pricing philosophies, such as premium or discount positioning. If these vendors have unique market positions that attract particular types of consumers, the model may reflect not only price dynamics but also underlying consumer segmentation. This brand-driven bias could distort interpretations, particularly if certain vendors consistently maintain higher or lower prices due to factors outside competitive dynamics (e.g., brand loyalty or perceived product quality).

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

\newpage

# References